---
title: "Personal task"
author: "Nazar Kalinichenko"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(xts)
```

## Content

I have included a file containing 5 years of stock data (in the allstocks5yr.csv).

The allstocks5yr.csv contains the merged data for individual stocks, presented in a merged .csv file.

The file has the following columns:
Date - in format: yy-mm-dd

Open - price of the stock at market open (this is NYSE data so all in USD)

High - Highest price reached in the day

Low Close - Lowest price reached in the day

Volume - Number of shares traded, pcs.

Name - the stock's ticker name

### Credit
https://www.kaggle.com/datasets/camnugent/sandp500

## Data reading
```{r}
setwd("D:/Projects/r/knu-pzod/Personal Task")
data <- read.csv("all_stocks_5yr.csv")
head(data)
tail(data)
```

## Data tidying
Get monthly IBM and AMD data:

```{r}
ibm_data <- data[data$Name == "IBM",]
rownames(ibm_data) <- as.Date(ibm_data$date)
ibm_data <- subset(ibm_data, select = -c(date, Name) )
head(ibm_data)
ibm_data_weekly <- apply.weekly(ibm_data, FUN = mean)
head(ibm_data_weekly)
ibm_data_monthly <- apply.monthly(ibm_data_weekly, FUN = mean)
head(ibm_data_monthly)

amd_data <- data[data$Name == "AMD",]
rownames(amd_data) <- as.Date(amd_data$date)
amd_data <- subset(amd_data, select = -c(date, Name) )
head(amd_data)
amd_data_weekly <- apply.weekly(amd_data, FUN = mean)
head(amd_data_weekly)
amd_data_monthly <- apply.monthly(amd_data_weekly, FUN = mean)
head(amd_data_monthly)
```

## Data analyzing with graphics
Get mean, standard deviation and variance of the stock prices at close:

```{r}
mean(ibm_data_monthly$close)
sd(ibm_data_monthly$close)
var(ibm_data_monthly$close)

mean(amd_data_monthly$close)
sd(amd_data_monthly$close)
var(amd_data_monthly$close)
```

Plot stock prices:
```{r}
plot.ts(ibm_data_monthly)
plot.ts(amd_data_monthly)
```

Plot stock percentage change:
```{r}
ibm_pct <- -diff(ibm_data_monthly$close)/ibm_data_monthly$close[-1]
amd_pct <- -diff(amd_data_monthly$close)/amd_data_monthly$close[-1]

plot.ts(ibm_pct * 100)
plot.ts(amd_pct * 100)
```

Plot stock cumulative product:
```{r}
ibm_cumprod <- cumprod(ibm_pct + 1)
amd_cumprod <- cumprod(amd_pct + 1)

plot.ts(ibm_cumprod)
plot.ts(amd_cumprod)
```